Development of D-box peptides to inhibit the Anaphase Promoting Complex/Cyclosome

Version 1

Abstract

E3 ubiquitin ligases engage their substrates via ‘degrons’ - short linear motifs typically located within intrinsically disordered regions of substrates. As these enzymes are large, multi-subunit complexes that generally lack natural small-molecule ligands and are hard to drug via conventional means, alternative strategies are needed to target them in diseases, and peptide-based inhibitors derived from degrons represent a promising approach. Here we explore peptide inhibitors of Cdc20, a substrate-recognition subunit and activator of the E3 ubiquitin ligase the anaphase promoting complex/cyclosome (APC/C) that is essential in mitosis and consequently of interest as an anti-cancer target. APC/C engages substrates via degrons that include the ‘Destruction box’ (D-box) motif. We used a rational design approach to construct binders containing unnatural amino acids aimed at better filling a hydrophobic pocket on the surface of Cdc20. We confirmed binding by thermal-shift assays and surface plasmon resonance and determined the structures of a number of the Cdc20-peptide complexes. Using a cellular thermal shift assay we confirmed that the D-box peptides also bind to and stabilise Cdc20 in the cell. We found that the D-box peptides inhibit ubiquitination activity of APC/CCdc20 and are more potent than the small molecule inhibitor Apcin. Lastly, these peptides function as portable degrons capable of driving the degradation of a fused fluorescent protein. Interestingly, we find that although inhibitory activity of the peptides correlates with Cdc20-binding affinity, degradation efficacy does not, which may be due to the complex nature of APC/C regulation and effects of degron binding of subunit recruitment and conformational changes. Our study lays the groundwork for the further development of these peptides as molecular therapeutics for blocking APC/C as well as potentially also for harnessing APC/C for targeted protein degradation.

Introduction

Interactions between E3 ubiquitin ligases and their substrates occur through “degrons” - short linear motifs (SLiMs) typically located within intrinsically disordered regions (IDRs) of substrates and which have relatively weak, micromolar affinities for the E3 (Min, Mayor, and Lindon 2013; Guharoy et al. 2016). Degradation of a substrate requires extensive sequential ubiquitination of the substrate through multiple rounds of recruitment of a ubiquitin-loaded conjugating E2 enzyme to the E3-substrate complex, and proteasome recognition depends on the number, type and length of polyubiquitin chains. In many examples explored to date, target discrimination and productive complex formation (i.e., leading to degradation of substrate) relies on the interaction of the E3 with multiple degrons ((Karamysheva et al. 2009; Fiore et al. 2016; Pierce et al. 2016; Tian et al. 2012; Okoye et al. 2022), and there is often a correlation between E3-substrate affinity and rate and/or timing of degradation.

Progression of cells out of mitosis involves ordered ubiquitin-mediated destruction of at least 100 different protein targets under control of a large multi-subunit E3 ubiquitin ligase, the anaphase promoting complex/cyclosome (APC/C), bound to one of its two coactivators Cdc20 and Cdh1/FZR1 (Davey and Morgan 2016; Min, Mayor, and Lindon 2013; Bakos et al. 2018). These activators contain a WD40 domain that binds substrate degrons flanked by partially disordered regions that mediate binding to neighbouring APC/C subunits, resulting in conformational change that enhances binding of the E2. APC/CCdc20 is a key component of the cell cycle machinery, with full activation of APC/CCdc20 acting as the trigger for mitotic exit through targeted degradation of mitotic cyclins and securin (PTTG1) (Meadows and Millar 2015). Coordination of mitotic exit events with segregation of duplicated chromosomes requires careful control of APC/CCdc20 activity, which is achieved via the mitotic checkpoint that inhibits APC/C in the presence of faulty chromosome attachments to the mitotic spindle (Izawa and Pines 2011; Hein and Nilsson 2014; Fiore et al. 2016; Qiao et al. 2016; Alfieri, Zhang, and Barford 2017; Watson et al. 2019). The mitotic checkpoint complex (MCC) prevents APC/CCdc20 targeting of critical metaphase substrates, whilst allowing degradation of a small number of so-called “checkpoint-independent” substrates such as cyclin A2 and Nek2A (Geley et al. 2001; Hayes et al. 2006). As cells exit mitosis, APC/CCdc20 activity declines and is replaced with APC/CFZR1, which maintains activity until the end of G1 phase.

In the past 15 years, high-resolution X-ray and cryo-EM (electron microscopy) studies of the APC/C and its interactions with substrates and E2s has generated a detailed description of the structure-function relationships that drive ubiquitination and degradation (Barford 2020). The binding of Cdc20 or FZR1 to the core APC/C creates at least three degron-binding sites for the known APC/C degrons, namely the “Destruction-box” (D-box, consensus RxxLxxxxN) and KEN motifs, and the ABBA motif thought to be required for Cyclin A degradation only(Qin et al. 2017). A cryo-EM study of the structure of APC/C-FZR1 in complex with its pseudo-substrate inhibitor Acm1 revealed simultaneous engagement of D-box, KEN, and ABBA motifs of Acm1 with their respective receptor sites on the interactions (He et al. 2013). The KEN motif docks to the top surface of the WD40 propeller of the co-activator and the D-box to a cleft formed between two blades of the propeller and the neighbouring APC10 subunit such that substrate engagement with degron receptors is likely to stabilize the active complex (Burton and Tsakraklides 2005; Buschhorn et al. 2010; Da Fonseca et al. 2010; L. Chang et al. 2014; Matyskiela and Morgan 2009; Qin et al. 2019). The critical residue of the D-box, leucine at position 4 (P4), contacts a hydrophobic pocket in the co-activator subunit, and the ‘tail’ of the D-box degron and its flanking sequence (P8-12) contact the APC10 subunit. Mechanisms by which the APC/C “orders” the degradation of its substrates include co-activator switching, fine-tuning of APC/C-substrate interactions by phosphorylation or other post-translational modifications, differential processivity of ubiquitination, and substrate competition, in addition to differential degron-binding affinities (Davey and Morgan 2016; Alfieri, Zhang, and Barford 2017; Bodrug et al. 2021; Okoye et al. 2022). Most recently, single-molecule studies have shed new insights into the key role of degron multivalency in enabling efficient substrate ubiquitination and degradation (Hartooni et al. 2022).

Inhibitors of APC/CCdc20 activity represent an interesting therapeutic approach to target dividing cells in cancer. Given the large size of the APC/C machine (11 subunits) and the complex mechanisms described above that regulate its function, it is not surprising that it is challenging to target. Apcin and TAME are recently identified small-molecule inhibitors, but they have limited activity and complicated output (Richeson et al. 2020; Sackton et al. 2014a). In this paper we use a rational approach, based on D-box consensus sequences and a ‘Super D-box’ peptide derived from Hsl1, and examination of the Cdc20-degron interface, to design a series of more potent binders containing unnatural amino acids aimed at better filling the hydrophobic pocket on the interaction interface. We quantified binding by thermal shift assays (TSA) and surface plasmon resonance (SPR) and used a cellular thermal shift assay (CETSA) to demonstrate target engagement within the cellular context. The peptides also show functional engagement with APC/C in the cell as evidenced by their ability to drive the degradation of a fluorescent protein. Most strikingly, in vitro ubiquitination assays with recombinant APC/CCdc20 shows that these peptides are more potent inhibitors of Cyclin B1 ubiquitination than Apcin. Interestingly, we find that although inhibitory activity of the peptides correlates with Cdc20-binding affinity, their degradation efficacy does not. This may be due to the complex nature of APC/C degrons and their bipartite interaction with different subunit, role in E2 recruitment, and consequent impact of positioning for effective ubiquitination. The results are a useful starting point for the further development of these peptides as molecular therapeutics for blocking APC/C as well as potentially also for harnessing APC/C for targeted protein degradation.

Materials and Methods

Cloning, expression, and purification of Cdc20

DNA encoding residues 161-477 of human Cdc20 (Cdc20WD40) was cloned into a pU1 vector with an N-terminal His6-tag followed by a TEV protease cleavage site. Plasmid was then transformed into DH10 MultiBac cells expressing a Cre-recombinase. Positive clones were grown up and bacmid DNA prepared by standard protocols. Sf21 or Sf9 cells were grown at 27°C in Erlenmeyer flasks (Corning) and maintained in mid-log phase of growth prior to all experiments. High titre baculovirus was produced by transfecting bacmid DNA into Sf21 cells at 0.5 × 106 cells/ ml cells using Superfect (Qiagen) in 24 deep-well blocks. Virus was harvested 1-week post transfection. For protein over-expression, Sf21 or Sf9 cells were infected with the virus stock and harvested about 60 hours post infection. Cell pellets were resuspended in 50 mM Tris-HCl, 300 mM NaCl, 1 mM MgCl2, 1 mM TCEP, 5% (v/v) glycerol, SigmaFAST EDTA-free protease inhibitor cocktail (1 tablet/100 ml), pH 8.5. Resuspended pellets were lysed by one freeze-thaw cycle at -80°C. Lysates were then clarified by centrifugation at 45,000 × g for 45 minutes at 4°C. Supernatants were flowed over a 5 ml HisTrap Excel column and washed with 20 column volumes (CV) of 50 mM Tris-HCl, 300 mM NaCl, 10 mM imidazole, 1 mM MgCl2, 1 mM TCEP, 5% (v/v) glycerol, pH 8.5. Proteins were then eluted with the above buffer including 300 mM imidazole directly into a 26/10 desalting column pre-equilibrated in the above buffer without imidazole. Eluted protein fractions were then pooled, and the His6-tag was removed using His6-TEV protease (S219V) overnight at 4°C. Proteins were then flowed back over a 5 ml HisTrap Excel column, collecting the flow-through containing the non-tagged Cdc20 protein. Protein eluent was then diluted in 25 mM Tris-HCl buffer, 1 mM MgCl2, 1 mM TCEP, 5% (v/v) glycerol, pH 8.5 to a final concentration of 30 mM NaCl. Proteins were then loaded onto a MonoQ 10/100 GL column and eluted over 20 CV with 1 M NaCl. Protein fractions containing the Cdc20 protein were then pooled and concentrated before separating on a Superdex 75 increase 10/300 GL column in the final buffer containing 25 mM Tris-HCl, 150 mM NaCl, 1 mM MgCl2, 5 mM TCEP, 5% (v/v) glycerol, pH 8.5.

Minimal biotinylation of Cdc20 for SPR

The protocol for minimal biotinylation was adapted from Papalia and Myszka (Papalia and Myszka 2010). Purified Cdc20WD40 was diluted into the reaction buffer (10 mM Tris-HCl, 150 mM NaCl, 1 mM MgCl2, 5% (v/v) glycerol, 1 mM TCEP, pH 7.3). A 0.9:1 molar ratio of Sulfo-NHS-LC-LC-Biotin (Thermo Fisher Scientific, A35358) was added to the diluted Cdc20 protein. The contents were briefly mixed by vortex and incubated on ice for 3 hours. The sample was then separated on a Superdex 75 Increase 10/300 GL column to remove free biotin.

Peptide synthesis and purification

Peptides synthesis was performed on a 0.1 mmol scale using Ramage-ChemMatrix® resin (Sigma Aldrich). Fmoc-L-amino acids (2 eq.), HATU (2 eq.) and HOAt (2 eq.) were dissolved in 2 mL of NMP. DIEA (3.4 eq.) were used to activate the coupling mixture. Activated Fmoc-L-amino acids were coupled for 10 minutes (Fmoc-L-Arginine, 2 × 5 eq., 30 minutes). Resins were washed in DMF and deprotected in 20% piperidine in DMF for 15 minutes. All peptides were N-terminally acetylated in 4 ml DMF, 4 mL acetic anhydride, 2 mL DIEA for 10 minutes. A peptide cleavage cocktail consisting of 93% TFA, 3.5% TIPS and 3.5% ddH2O was used to deprotect and cleave the peptide from the resin for 1 hour. The eluate was triturated by the addition of diethyl ether and the resulting precipitate was isolated by brief centrifugation. All peptides were characterised by LCMS using a Waters ACQUITY H-Class UPLC with an ESCi Multi-Mode Ionisation Waters SQ Detector 2 spectrometer. LC was performed on a ACQUITY UPLC® CSH C18 (2.1 mm × 50 mm, 1.7 µm, 130 Å) at 40°C, with a PDA eλ detector 220 – 800 nm, interval 1.2 nm. The following solvents and gradients were used for LC runs. Solvent A: 2 mM NH4OAc in 95% H2O, 5% MeCN, solvent B: 100% MeCN, solvent C: 2% Formic acid from 5-95% B with a constant of 5% C over 1 minute at 0.6 ml/min. Analytical and semi preparative HPLC runs were performed on an Agilent 1260 Infinity system using a Supelcosil ABZ+PLUS (150 mm × 4.6 mm, 3 µm) and Supelcosil ABZ+PLUS (250 mm × 21.2 mm, 5 µm), respectively. Peptides were eluted with a linear gradient system (solvent A: 0.1% TFA in H2O, solvent B: 0.05% TFA in MeCN) over 15 minutes at 1ml/min and 20 minutes at 20ml/min, respectively. Eluents were monitored by UV absorbance at 220 nm and 254 nm. Analytical data for all peptides are shown in Figure S5.

Thermal-shift assays (TSA)

Assays were performed using a Roche LightCycler 480 I in 96-well plate format. Each well (20 µl) was prepared with 750 nM of purified Cdc20WD40 and varying concentrations of D-box peptides, Apcin or DMSO (vehicle control) in assay buffer; 25 mM Tris-HCl, 150 mM NaCl, 1 mM MgCl2, 5% (v/v) glycerol, 1 mM TCEP, 1% (v/v) DMSO, 5x SYPRO Orange (Thermo Fisher), pH 8.5. Thermal ramps were conducted from 25°C to 95°C at a rate of 0.03°C/sec and data were collected at a frequency of 20 points/°C. An excitation wavelength of 483 ± 35 nm was used to excite SYPRO Orange, and the fluorescence emission was detected at a 568 ± 20 nm. Measurements were performed in triplicate and errors listed are the standard deviation. Melting temperatures were determined by the minima peak of the negative differential in the ‘Tm calling’ analysis within the in-built analysis software.

Surface Plasmon Resonance (SPR) assays

Experiments were performed using a Biacore T200 instrument (GE healthcare) at 15°C. Biotinylated-Cdc20WD40 was immobilised onto a SA biosensor chip (GE healthcare) in running buffer (10 mM HEPES, 150 mM NaCl, 0.1 mM TCEP, 0.05% (v/v) Tween 20 and 1% (v/v) DMSO, pH 7.4) over flow cells 2, 3 and 4 at varying ligand densities. Flow cell 1 was used as a reference cell. Free biotin binding sites were blocked using amine-PEG4-Biotin. Peptides and Apcin analytes were diluted from DMSO stock solutions in running buffer without DMSO and were buffer matched to 1% DMSO. Titrations of each analyte were run over the sensor chip at a flow rate of 30 µl/min. Binding interactions were detected as a change in response units over the reference flow cell and subtracted from a blank buffer injection. Dissociation constants (KD) were calculated by fitting the response units (RU) at steady-state equilibrium generated by the binding of an analyte to Cdc20WD40 against the concentration of analyte using the following equation: where RUanalyte is the response units at equilibrium during a given injection of a concentration of analyte, [analyte]. RUmax is the maximum response produced by the a given analyte, dependant on the RU of immobilised ligand on a given flow cell. KD is the dissociation constant of a given analyte to the ligand. KD values are shown as the average of measurements from the three reference-subtracted flow cells.

Cellular thermal shift assays (CETSA)

Full-length Cdc20 (residues 1-499) with a C-terminal HiBiT tag (GSVSGWRLFKKISGS, Promega) was cloned into a pcDNA3.1(-) vector. HEK 293T cells were cultured in DMEM + 10% FBS (Sigma Aldrich, F7524) at 5% CO2 in a humidified environment. Cells were grown to 70% confluency in T75 flasks prior to transient transfection with 10 µg of Cdc20_HiBiT_pcDNA3.1(-) plasmid with Lipofectamine 2000 (Invitrogen, ThemoFisher Scientific) according to the manufacturers’ protocol. Cells were harvested after 48 hours by trypsinisation and were subsequently washed twice in PBS with repeated centrifugation at 1000 × g for 2 minutes. The pellet was then resuspended in lysis buffer (PBS, 1 × SigmaFAST EDTA-free protease inhibitor tablet (Sigma Aldrich), 2 mM NaVO3, 5 mM NaF, pH 7.4) and freeze-thaw lysed in liquid nitrogen. The lysate was clarified by centrifugation at 20,000 × g, 4°C for 20 minutes and the protein concentration of the supernatant was quantified by BCA (Pierce). Lysates were used at a final concentration of 0.2 mg/ml in lysis buffer. Lysates were aliquoted in 300 µL and were spiked with D-box binding site ligands to a concentration of 100 µM maintaining 1% DMSO. Compounds were incubated on ice for 30 minutes prior to aliquoting further into PCR strip tubes on a PCR block at 4°C. Lysate aliquots were then heated on a second PCR block at the indicated temperatures for 3 minutes prior to returning to 4°C. Heated lysates (5 µl) were then transferred into an AlphaPlate light-grey 384-well plate in quadruplicate by multichannel pipette. Nano-Glo HiBiT lytic detection system (Promega) was diluted as per the manufacturers’ instructions and 5 µl were added to each well by multichannel pipette. Lysis buffer and a non-transfected HEK 293T cell lysate were used as negative controls. Following five minutes of incubation on a plate shaker, the plate was measured using a CLARIOStar microplate reader (BMG Labtech), with the detector set to read at 460 ± 80 nm, the focal height at 10.5 cm and the gain adjusted to 2000. Data were normalised to the unheated sample (4°C) and were fitted using a Boltzmann equation to extract the melting temperature (Tm) (Niesen, Berglund, and Vedadi 2007).

Protein crystallisation

Peptide D21 was added to Cdc20WD40 in a stoichiometric manner and was co-concentrated to 1.9 mg/ml. The resulting complex was crystallised in a 2:1 protein to well solution ratio at 20°C using the sitting-drop vapour-diffusion method with a well solution containing 0.1 M MES pH 6.5, 12% (w/v) PEG 6000, 10% (v/v) MPD for Cdc20WD40-D21 and 0.1 M MES pH 6.5, 14% (w/v) PEG 6000, 10% (v/v) MPD for Cdc20WD40-D20 and Cdc20WD40-D7. Crystals grew to a maximum size after 3 days of incubation. For soaking experiments, crystals were first looped and washed through three drops containing 0.1 M MES pH 6.5, 20% (w/v) PEG 6000 to wash out MPD from the crystal. Crystals were then looped and incubated in a solution containing 0.1 M MES pH 6.5, 20% (w/v) PEG 6000 and 2.5 mM D21 or D20 (5% (v/v) DMSO) or D7 (10% (v/v) DMSO) for four hours. Soaked crystals were cryo-protected in the soak solution supplemented with 10% (v/v) glycerol and were flash frozen in liquid nitrogen.

Data collection and structure determination

Diffraction data were collected on beamline I04 at the Diamond Light Source (Oxford, UK) and processed using autoPROC-STARANISO STARANISO (Vonrhein et al. 2018). Phases were obtained by molecular replacement using the crystal structure of human Cdc20 (PDB ID code 4GGC) as the search model (Tian et al. 2012). Iterative model building and refinements were performed with COOT and BUSTER, respectively (Emsley 2010, Bricogne G. et al) Cdc20-D20 and Cdc20-D21 datasets were first refined using Refmac5 within the CCP4i suite ((Winn et al. 2011; Kovalevskiy et al. 2018) before final refinements using BUSTER. Data collection and structure refinement statistics are summarised in Table S1.

Ubiquitination assays

In vitro ubiquitination experiments were performed using APC/C and Cdc20 purified from insect cells (Zhang et al. 2016). 60 nM APC/C, 30 nM Cdc20, 90 nM UBA1, 300 nM UbcH10, 300 nM Ube2S, 35 mM ubiquitin, 1 mM cyclin B1, 5 mM ATP, 10 mM MgCl2, were mixed in a buffer containing 40 mM HEPES (pH 8.0), 80 mM NaCl, 0.6 mM DTT. The reaction was either performed with the indicated concentrations of peptides or DMSO (Sigma Aldrich) as the vehicle control. The reaction was incubated for 30 min at 23 °C and stopped by the addition of one volume of 2x concentrated NuPAGE LDS loading buffer (Invitrogen).

Protein degradation assays

The pEGFP-N1 vector was modified by swapping the EGFP-coding sequence for mNeon-coding sequence using the AgeI/NotI cloning sites. The Aurora kinase A (AURKA) C-terminal fragment (364-403) containing the non-degron R371xxL motif (D0) together with an extended IDR was amplified by PCR and cloned into the modified vector with BamHI/AgeI sites. Round the horn site-directed mutagenesis was used to generate different D-box variants and validated by DNA sequencing. U2OS cells were cultured in DMEM supplemented with 10% FBS, 200 µM Glutamax-1, 100 U/ml penicillin, 100 µg/ml streptomycin, and 250 ng/ml fungizone (all from ThermoFisher Scientific) at 37ºC in humidified atmosphere containing 5% CO2. Plasmids were introduced into U2OS cells by electroporation using the Neon™ Transfection System 10 μL Kit (ThermoFisher Scientific) and cells seeded on eight-well microscopy slides (Ibidi) and recovered for 24 hours. DMEM medium was exchanged for phenol red-free Leibovitz’s L15 (ThermoFisher Scientific), supplemented as above. Time-lapse imaging was conducted at 37°C using a widefield imaging platform composed of Olympus IX83 motorized inverted microscope, Spectra-X multi-channel LED widefield illuminator (Lumencor, Beaverton, OR, USA), Optospin filter wheel (Cairn Research, Faversham, UK), CoolSnap MYO CCD camera (Photometrics, Tucson, AZ, USA), automated XY stage (ASI, Eugene, OR, USA) and climate chamber (Digital Pixel, Brighton, UK), all controlled using Micro-Manager software (Edelstein et al. 2014). Fluorescence and phase contrast images of cells in mitosis were acquired with a 40X objective binned at 2x2 at 2-minute intervals. Fluorescence intensity of mNeon in individual mitotic cells was quantified from 16-bit tiff files using ImageJ, by integrating pixel measurements after subtraction of background fluorescence. Degradation curves were synchronized in silico to anaphase onset to generate average curves for multiple cells in each experimental condition.

Results

Quantification of Cdc20-binding activity of the small molecule Apcin

We first produced Cdc20 protein in sufficient quantities for biophysical analysis and then used the known small molecule binder, Apcin, to test that the purified protein was functional and to benchmark our peptide-binding measurements. As Cdc20 comprises a WD40 domain that binds to the different degrons and is flanked on each end by long intrinsically disordered regions, we made a construct comprising the WD40 domain (residues 161 to 477) with an N-terminal His6-tag and expressed this protein in baculovirus as previously described (Sackton et al. 2014b; Tian et al. 2012). We biotinylated Cdc20 at a single-site, as shown by electrospray-ionisation mass spectrometry (Fig. S1). Using TSA and SPR, we confirmed that the purified Cdc20 was capable of binding to Apcin. The KD obtained by SPR was 420 ± 50 nM (Fig. 2).

Structure and Function of APC/C.

(A) Schematic of APC/C activity during mitotic exit, indicating the switch in co-activator from Cdc20 to FZR1. Most substrates contain variable degrons (D-box in green, KEN in yellow) present in IDRs (B) Domain structuring of Cdc20 comprising an N-terminal IDR with the C-box, KEN-box, and CRY-box motifs, the central WD40 domain responsible for substrate recruitment via the degron binding sites and the C-terminal IDR containing the IR-tail. (C) Schematic of the structure of the Cdc20 WD40 domain (PDB: 4GGC) overlaid with those of the WD40 domain in complex with Acm1 D-box and ABBA motif peptides (PDB: 4BH6) (He et al. 2013) and the KEN-box peptide 4GGD) (Tian et al. 2012).

Biophysical characterisation of Apcin binding to Cdc20WD40 by TSA and SPR.

(A) Representative examples of thermal unfolding traces of Cdc20 WD40 in the presence of 1% DMSO as the vehicle control or Apcin at concentrations of 25, 50 and 100 µM. (B) Corresponding melting temperatures calculated from derivative plots of the thermal unfolding traces. Mean data from triplicate measurements are shown, with error bars representing standard deviations. (C) Reference-subtracted sensorgrams of biotinylated Cdc20WD40 and Apcin. (D) Binding affinity determination of Apcin to Cdc20WD40 domain by steady-state analysis of the sensorgrams.

Design of D-box peptides

Structures of the D-box-APC/C interactions [(Chao et al. 2012; He et al. 2013)], which show that there are three key residues, Arginine 1, Leucine 4 and Asparagine 9 of the RxxLxxxxN motif (Fig. 3A). As a starting point we used two peptides, a 10-residue consensus-like sequence derived from Hsl1 (D1: GRAALSDITN) (Burton and Tsakraklides 2005; Frye et al. 2013; Davey and Morgan 2016), and a 9-residue consensus sequence based on known D-box degrons from APC/C substrates (D2: RLPLGDISN) (He et al. 2013). TSA and SPR showed that D1 binds to Cdc20WD40 with a weak affinity (KD = 18.6 ± 0.2 µM) (Table 1). D2 had no detectable affinity by TSA and was consequently not analysed by SPR. We hypothesised that the apparent lack of binding may be due the low solubility in aqueous buffer of D2 rather than an inability to bind. Based on the Cdh1-Acm1 structure, the sidechain of the amino acid at position 2 is likely to be solvent exposed in the context of Cdc20WD40 (Fig. 3A). A substitution was therefore made at this position, from Leu to Ala, which improved peptide solubility and was therefore incorporated in all subsequent peptides.

D-box peptide mutations.

(A) Schematic showing the Acm1 D-box peptide bound to yeast FZR1 homologue Cdh1. R119 of the D-box forms H-bond interactions with D256 and E537 of Cdh1. L122 of the D-box buries into the canonical pocket on the surface of Cdh1 (PDB: 4BH6, He et al. 2013). (B) Melting temperature of Cdc20WD40 in the presence of D-box peptides at 25, 50 and 100 µM concentrations, calculated from derivative plots of the thermal unfolding traces. Mean data from triplicate measurements are shown, with error bars representing standard deviations.

Isoleucine at position 7 and Proline at position 3 of the D-box peptide are optimal for binding

From the consensus sequence, we observed that substrate proteins have approximately equal frequency of Val, Leu and Ile at position 7. Based on the yeast Cdh1-Acm1 X-ray crystal structure (Fig. 3A), this interaction appears atypical of hydrophobic interactions given the largely solvent-exposed nature of the amino acid sidechain. Given the similar structural and physical properties of the three aliphatic sidechains, we compared peptides with each of these three amino acids at position 7 and found that D4 with Ile7 had the highest affinity for Cdc20 (1.5-fold higher than D5 with Leu7 (19.6 ± 0.2 µM and 27 ± 1 µM respectively) (Table 1 and Fig. S2C, D). Interestingly, the shorter hydrocarbon chain of Val in D3 gave the weakest affinity, with a KD determined by SPR at 54.4 ± 0.7 µM.

We next investigated the contribution of Proline versus Alanine at position 3 (Table 1). Like the position 7 residues, Pro and Ala appear in approximately equal distribution to each other among known substrate proteins. In the context of D-box degron binding, modelling of our D4 peptide to the S. cerevisiae Cdh1 structure showed that Pro 3 may form a favourable turn in the D-box peptide backbone to allow the side chain of Leu 4 to adopt its canonical pocket (Fig. 3A). To test this hypothesis, we proceeded to synthesise D10, a derivative of D4 containing an A3P single point mutation. As expected, this mutation was significantly detrimental with an affinity of 70 ± 3 µM by SPR and in parallel a loss of thermal stabilisation by TSA (Fig. 3B). Upon confirming our hypothesis, we synthesised a derivative of D1 containing the A3P point mutation, yielding D19 (RAPLSDITN). This substitution resulted in 3-fold increase in affinity (KD = 5.9 ± 0.1 µM) compared with its parental sequence (Fig. S2F, Table 1).

Unnatural amino acids at position 4 of the D-box peptide result in significantly enhanced binding affinity to Cdc20

The surface topology of Cdc20 is largely flat, making it hard to drug. Nevertheless, in Apcin the tri-chlorinated moiety makes particular use of the Leu 4-binding pocket on Cdc20. Taking inspiration from the small molecule, we explored candidate unnatural amino acids to incorporate into the D-box peptides at position 4. Given that the pocket can accommodate a tri-chlorinated carbon moiety within Apcin, we explored similar moieties to append to our D-box peptides. We incorporated (S)-2-amino-4,4-dimethylpentanoic acid (C3) (Fig. 4A) into the backbone sequences of D4, D10 and D19 replacing Leu at position 4, yielding peptides D7, D12, and D20, respectively (Table 2). As expected, the structure-activity relationship (SAR) held true between all peptides, whereby incorporation of the unnatural amino acid increased the binding affinity over 6-fold versus the respective parental peptide (Table 2). Building on this success, we further explored the commercially available halogenated analog, (S)-2-amino-4,4,4-trifluorobutanoic acid (F3) (Fig. 4A), leading to peptide D21 (Table 2). With the tri-fluorinated group, a further increase in binding affinity was achieved (KD = 520 ± 10 nM), which is similar to that of Apcin.

D-box peptides incorporating unnatural amino acids.

(A) Schematics of the two unnatural amino acids used. (B) Thermal stabilisation of the Cdc20WD40 by the two highest affinity peptides D20 and D21 calculated from derivative plots in TSA. SPR reference-subtracted sensorgrams and binding curves for (C) D20, and (D) D21.

Crystal structures of Cdc20-peptide complexes reveal D-box binding mode

Previous attempts to co-crystallise Cdc20 and securin-derived or cyclin B1-derived D-box peptides by Tian and co-workers were unsuccessful (Tian et al. 2012), which may be due to the low affinity of peptides comprising these sequences. Despite the relatively high affinity of D21 and the approximate 1:1.5 ratio of protein to peptide used in co-crystallisation experiments, crystals were absent of peptide ligands and instead contained the 2-methyl-2,4-pentanediol (MPD) molecule in the Leucine-binding cleft (data not shown), originating from the crystallisation well solution. We therefore adopted a similar protocol to that described by Sackton et al., whereby MPD was ‘washed’ out from the crystal prior to performing a soaking experiment with the desired ligand. We attempted these soaking experiments with our four highest affinity peptides, D21, D20, D7 and D19 (in order of highest affinity to lowest) and were able to observe sufficient ligand density for all but D19.

The crystal structures of Cdc20WD40 in complex with each of the other three D-box peptides (Fig. 5A-C) show that they bound to Cdc20WD40 at the canonical D-box degron binding site, with a largely similar topology to the S. cerevisiae Acm1-Cdh1 structure (Fig. 5D (overlay of D21 with Acm1 D-box). The R1 guanidino group of peptides interacts forms hydrogen bonds with the carboxylic acid side chains of D177 and E465 of Cdc20WD40. The nitrogen backbone atom of the (S)-2-amino-4,4,4-trifluorobutanoic acid/ (S)-2-amino-4,4-dimethylpentanoic acid unnatural amino acids also form a hydrogen bond with the carbonyl of D177. Additionally, the carbonyl of S5 belonging to D21/D20 form a H-bond with the nitrogen backbone atom of D177. Lastly, D6 forms inter-molecular H-bonds with R174. We also observed intra-molecular H-bond between the carbonyl of A2 with the amine of G5/S5, in addition the carbonyl of A2 to the hydroxyl of S5 in D21/D20. Crystal packing of an adjacent asymmetric unit of the WD40 domain likely occludes the assumed binding site for the C-terminal three residues (…ITN-NH2). We therefore presume this is the reason for lack of observed density in this region of the peptides D20 and D21 (Fig. S3E and S3F, respectively). We extend this observation to further explain why we were unable to observe peptide density for the D19-soaked crystals. Specifically, our affinity data highlights a role of position 7 in binding, which in crystallo is unable to be realised. This hypothesis also correlates to the comments made by Tian et al. in their attempt to co-crystallize securin D-box peptides with Cdc20, in the identical space group (Tian et al. 2012).

Crystal structures of Cdc20-D-box complexes.

X-ray crystal structures of peptides (A) D21, (B) D20 & (C) D7 bound to the canonical D-box binding pocket of Cdc20. Intermolecular hydrogen bonds between peptides and Cdc20 are shown by dashed lines. (D) Structural alignment of D21-bound Cdc20 and Acm1 D-box peptide bound to Cdh1 (PDB: 4BH6 (He et al. 2013)). Peptide backbones align to with an RMSD of 1.007Å. Modelled water molecules have been removed from images for clarity.

D-box peptides bind to Cdc20 in the cellular context

We next investigated whether the four highest-affinity peptides D21, D20, D7, and D19 can bind to Cdc20 in the cellular context using a cellular thermal shift assay (CETSA) (Martinez Molina 2013). Sackton et. al previously demonstrated that Apcin can stabilise endogenous Cdc20 by using an isothermal CETSA method (Sackton et al. 2014b). We were able to reproduce this ligand-induced stabilisation of Cdc20 using the more commonly used temperature gradient approach by densitometric analysis of western blots (Fig S3A). However, due to the low-throughput of the assay we also explored a more high-throughput approach by making use of Promega’s split-luciferase HiBiT tag appended to the C-terminus of full-length Cdc20, based on protocols previously described by Martinez and co-workers (Martinez et al. 2018). Notably, the signal is more sensitive and has a larger range of compared to a western blot, and it removes a significant time-consuming centrifugation step from the workflow. We first confirmed that omitting the centrifugation step did not significantly affect the observed Tm of vehicle control samples (Fig. S4B). To further validate that the transfected Cdc20 is functional, we probed binding of 100 µM Apcin, which gave a Tm of 54.4°C ± 0.6 °C (Fig. S4C). We then explored whether the D-box peptides at a fixed concentration stabilise the Cdc20, and for D7, D20 and D21 we observed increases in the thermal stability of Cdc20 that correlated with their binding affinities as previously determined (Fig. 6 and Table 3). The lowest-affinity peptide, D19, did not result in a significant thermal stabilisation of Cdc20.

D-box peptides bind to full-length HiBiT-tagged Cdc20 in the cellular context.

Representative CETSA data are shown for Cdc20-tranfected HEK293T cell lysates incubated with D-box peptides at a concentration of 100 µM.

D-box peptides inhibit APC/CCdc20 ubiquitination activity

We next assessed whether D21 and D20, the two highest affinity peptides, are able to inhibit APC/CCdc20 activity. In the context of Cyclin B1 ubiquitination, we found that both peptides are more potent inhibitors compared with Apcin at the same concentration despite having slightly lower Cdc20-binding affinities than Apcin (Fig. 7).

Inhibition of APC/CCdc20-mediated ubiquitination of Cyclin B1 by D-box peptides and Apcin.

In vitro ubiquitination assays using reconstituted APC/CCdc20 with Cyclin B1 as the substrate for ubiquitination. Lead peptides and Apcin were titrated from 300 µM to 3 µM and showed concentration-dependent inhibition of Cyclin B1 ubiquitination compared to the vehicle control (0.7% DMSO).

D-box peptides are able to target mNeon for degradation

To probe the functionality of the D-box variants at the cellular level, we conducted live cell degradation assays using mNeon fusions containing those peptide sequences that contain only natural amino acids: D1, D2, D3, and D19 (Fig. 8A). The D-box sequences were swapped into an RxxL motif previously shown to have no degron activity (Abdelbaki et al. 2022) which we here refer to as ‘D0’, adjacent to the endogenous C-terminal IDR of AURKA to enable processing of the ubiquitinated fusion proteins at the 26S proteasome. We found that all four new D-box variants tested could target mNeon for degradation, with timing consistent with targeting by APC/CCdc20 (Fig. 8B). We predicted that the higher affinity D-box peptides from the in vitro assays (D1 and D19) would mediate increased rates and extent of degradation compared to the lower affinity peptides (D2 and D3). However, we found the opposite effect: D2 and D3 showed increased rates of mNeon degradation compared to D1 and D19 (Fig. 8C,D). This observation is consistent with the idea that high-affinity binding at degron binding sites on APC/C, such as in the case of the yeast ‘pseudo-substrate’ inhibitor Acm1, acts to impede polyubiquitination of the bound protein (Qin et al. 2019). Indeed, there is no evidence that Hsl1, which is the highest affinity natural D-box (D1) used in our study, is degraded any more rapidly than other substrates of APC/C in yeast mitosis. As shown in Qin et al., mutation of the high affinity D-box in Acm1 converts it from inhibitor to substrate (Qin et al. 2019). Overall, our results support the conclusions that all the D-box peptides engage productively with the APC/C and that the highest affinity interactors act as inhibitors rather than functional degrons of APC/C.

D-box variants can drive degradation in mitotic cells.

(A) Schematic of D-box-mNeon constructs used in fluorescence timelapse imaging. (B) mNeon fluorescence levels in individual cells plotted over time to show D-box mediated degradation of mNeon in mitosis. Fluorescence measurements from individual cells are normalized to fluorescence at metaphase then in silico synchronized to anaphase onset. Mean degradation curves are shown, with error bars representing SDs. (C) Degradation rate curves show rate of change in relative fluorescence of the D-box variants and reveal maximum degradation rate for each construct. Error bars are depicted as shaded regions and indicate SDs. (D) Levels of relative fluorescence in each cell at t = 1 hour after anaphase onset. Degradation of each D-box construct was significant relative to D0 control, using Welch’s t-test. ****, p ≤ 0.0001. In (B)-(D), n = D0 (20) D1 (23), D2 (40), D3 (38), D19 (34) with data pooled from two or more independent experiments.

Discussion

Here we quantified D-box peptide binding to Cdc20 and show that binding affinities can be enhanced by incorporating unnatural amino acids to better fill the hydrophobic pockets on the Cdc20 surface. We confirmed the success of this approach by determining X-ray crystal structures of Cdc20-peptide complexes. We showed target engagement by the peptides in the cellular context, and we found that the two highest affinity peptides were more potent inhibitors of APC/CCdc20 activity than the small molecule Apcin. Lastly, we found that the D-box peptide is a portable motif that can drive productive ubiquitination leading to degradation when fused to a fluorescent protein target.

The finding that the peptides were more potent than Apcin as APC/CCdc20 inhibitors was somewhat surprising, since Apcin has a slightly higher Cdc20-binding affinity than the peptides. It suggests that inhibiting APC/CCdc20 ubiquitination activity may require larger molecules to compete with substrates effectively. It may also be that, unlike Apcin, the peptides not only block the interaction of substrates with Cdc20 but additionally the interaction with APC10 and/or prevent the conformational change in APC/C that enables recruitment of the E2. In fact, the mechanism of inhibition by Apcin and D-box peptides could be different – it may be that Apcin-bound Cdc20 can still bind to APC/C but peptide-bound Cdc20 cannot. Interestingly, although the inhibitory activity of the D-box peptides roughly correlates with the binding affinity, binding and degradation may be inversely correlated. In addition to the binding of D-box substrates to the co-activators Cdc20 and Cdh1, Qin et al have described how residues C-terminal of the D-box sequence, the ‘D-Box Extension’ DBE motif, influence recruitment of APC10 and potentially APC10 conformational changes enabling the recruitment of the E2 Ube2S (Qin et al. 2019). Ube2S is essential for adding K11 chains, and we showed previously that degradation of all substrates is dramatically slowed down by a lack of Ube2S (Min et al. 2015). The mNeon-D-box constructs used in our current study all contain the same DBE motif, so a potential contribution from this motif will not affect the interpretation of our results, but it could certainly be added as an element in future inhibitor design.

In summary, the finding presented here represent a useful starting point for the further development of APC/C inhibitors as both research tools and also molecular therapeutics. Future directions could involve enhancing potency through avidity by incorporating multiple degrons into our molecules and additions to the D-box core sequence to include motifs that engage other components of the APC/C machinery - namely APC10 and the E2 - thereby not only blocking substrate binding more effectively but also better impeding ubiquitination activity. The results also have implications for the design of small-molecule and peptide-based degraders that harness the APC/C.

eLife Assessment

The manuscript represents a fundamental advance in designing peptide inhibitors targeting Cdc20, a key activator and substrate-recognition subunit of the APC/C ubiquitin ligase. Supported by compelling biophysical and cellular evidence, the study lays a strong foundation for future developments in degron-based therapeutics. The unexpected findings regarding degradation efficiency highlight intriguing questions that merit further investigation. This work will interest researchers focused on peptide drug design targeting complex protein interactions.

Reviewer #1 (Public review):

Summary:

In this manuscript, the authors Eapen et al. investigated the peptide inhibitors of Cdc20. They applied a rational design approach, substituting residues found in the D-box consensus sequences to better align the peptides with the Cdc20-degron interface. In the process, the authors designed and tested a series of more potent binders, including ones that contain unnatural amino acids, and verified binding modes by elucidating the Cdc-20-peptide structures. The authors further showed that these peptides can engage with Cdc20 in the cellular context, and can inhibit APC/CCdc20 ubiquitination activity. Finally, the authors demonstrated that these peptides could be used as portable degron motifs that drive the degradation of a fused fluorescent protein.

Strengths:

This manuscript is clear and straightforward to follow. The investigation of different peptide variations was comprehensive and well-executed. This work provided the groundwork for the development of peptide drug modalities to inhibit degradation or apply peptides as portable motifs to achieve targeted degradation. Both of which are impactful.

Weaknesses:

A few minor comments:

(1) In my opinion, more attention to the solubility issue needs to be discussed and/or tested. On page 10, what is the solubility of D2 before a modification was made? The authors mentioned that position 2 is likely solvent exposed, it is not immediately clear to me why the mutation made was from one hydrophobic residue to another. What was the level of improvement in solubility? Are there any affinity data associated with the peptide that differ with D2 only at position 2?

(2) I'm not entirely convinced that the D19 density not observed in the crystal structure was due to crystal packing. This peptide is peculiar as it also did not induce any thermal stabilization of Cdc20 in the cellular thermal shift assay. Perhaps the binding of this peptide could be investigated in more detail (i.e., NMR?) Or at least more explanation could be provided.

Reviewer #2 (Public review):

Summary:

The authors took a well-characterised (partly by them), important E3 ligase, in the anaphase-promoting complex, and decided to design peptide inhibitors for it based on one of the known interacting motifs (called D-box) from its substrates. They incorporate unnatural amino acids to better occupy the interaction site, improve the binding affinity, and lay foundations for future therapeutics - maybe combining their findings with additional target sites.

Strengths:

The paper is mostly strengths - a logical progression of experiments, very well explained and carried out to a high standard. The authors use a carefully chosen variety of techniques (including X-ray crystallography, multiple binding analyses, and ubiquitination assays) to verify their findings - and they impressively achieve their goals by honing in on tight-binders.

Weaknesses:

Some things are not explained fully and it would be useful to have some clarification. Why did the authors decide to model their inhibitors on the D-box motif and not the other two SLiMs that they describe? What exactly do they mean when they say their 'observation is consistent with the idea that high-affinity binding at degron binding sites on APC/C, such as in the case of the yeast 'pseudo-substrate' inhibitor Acm1, acts to impede polyubiquitination of the bound protein'? It's an interesting thing to think about, and probably the paper they cite explains it more but I would like to know without having to find that other paper.

Reviewer #3 (Public review):

Summary:

Eapen and coworkers use a rational design approach to generate new peptide-inspired ligands at the D-box interface of cdc20. These new peptides serve as new starting points for blocking APC/C in the context of cancer, as well as manipulating APC/C for targeted protein degradation therapeutic approaches.

Strengths:

The characterization of new peptide-like ligands is generally solid and multifaceted, including binding assays, thermal stability enhancement in vitro and in cells, X-ray crystallography, and degradation assays.

Weaknesses:

One important finding of the study is that the strongest binders did not correlate with the fastest degradation in a cellular assay, but explanations for this behavior were not supported experimentally. Some minor issues regarding experimental replicates and details were also noted.

Author response:

The following is the authors’ response to the original reviews

Public reviews:

Reviewer #1 (Public review):

Summary:

In this manuscript, the authors Eapen et al. investigated the peptide inhibitors of Cdc20. They applied a rational design approach, substituting residues found in the D-box consensus sequences to better align the peptides with the Cdc20-degron interface. In the process, the authors designed and tested a series of more potent binders, including ones that contain unnatural amino acids, and verified binding modes by elucidating the Cdc-20-peptide structures. The authors further showed that these peptides can engage with Cdc20 in the cellular context, and can inhibit APC/CCdc20 ubiquitination activity. Finally, the authors demonstrated that these peptides could be used as portable degron motifs that drive the degradation of a fused fluorescent protein.

Strengths:

This manuscript is clear and straightforward to follow. The investigation of different peptide variations was comprehensive and well-executed. This work provided the groundwork for the development of peptide drug modalities to inhibit degradation or apply peptides as portable motifs to achieve targeted degradation. Both of which are impactful.

Weaknesses:

A few minor comments:

(1) In my opinion, more attention to the solubility issue needs to be discussed and/or tested. On page 10, what is the solubility of D2 before a modification was made? The authors mentioned that position 2 is likely solvent exposed, it is not immediately clear to me why the mutation made was from one hydrophobic residue to another. What was the level of improvement in solubility? Are there any affinity data associated with the peptide that differ with D2 only at position 2?

The reviewer is correct that we have not done any detailed solubility characterisation; we refer only to observations rather than quantitative analysis. We wrote that we reverted from Leu to Ala due to solubility - we have clarified this statement (page 11) to say that that we reverted to Ala, as it was the residue present in D1, for which we observed a measurable affinity by SPR and saw a concentration-dependent response in the thermal shift analysis. We do not have any peptides or affinity data that explore single-site mutations with the parental peptide of D2. D2 is included in the paper because of its link to the consensus D-box sequence and thus was the logical path to the investigations into positions 3 and 7 that come later in the manuscript.

(2) I'm not entirely convinced that the D19 density not observed in the crystal structure was due to crystal packing. This peptide is peculiar as it also did not induce any thermal stabilization of Cdc20 in the cellular thermal shift assay. Perhaps the binding of this peptide could be investigated in more detail (i.e., NMR?) Or at least more explanation could be provided.

This section has been clarified (page 16). The lack of observed density was likely due to the relatively low affinity of D19 and also to the lack of binding of the three C-terminal residues in the crystal, and consequently it has a further reduced affinity. The current wording in the manuscript puts greater emphasis on this second aspect being a D19-specific issue, even though it applies to all four soaked peptides. The extent of peptide-induced thermal stabilisations observed by TSA and CETSA is different, with the latter experiment consistently showing smaller shifts. This observation may be due to the more complex medium (cell lysate vs. purified protein) and/or different concentrations of the proteins in solution. In the CETSA, we over-expressed a HiBiT-tagged Cdc20, which is present in addition to any endogenously expressed Cdc20. Although we did not investigate it, the near identical D-box binding sites on Cdc20 and Cdh1 would suggest that there will be cross-specificity, which could further influence the CETSA experiments.

The section now reads:

“We therefore assume that this is the reason for the lack of observed density in this region of the peptides D20 and D21 (Fig. S3E and S3F, respectively). We believe that it causes a reduction in binding affinities of all peptides in crystallo, given the evidence from SPR highlighting a role of position 7 in the interaction (Table 1). Interestingly, the observed electron density of the peptide correlates with Cdc20 binding affinity: D21 and D20, having the highest affinities, display the clearest electron density allowing six amino acids to be modeled, whereas D7 shows relatively poor density permitting modelling of only four residues. For D19, the lack of density observed likely reflects its intrinsically weaker affinity compared to the other peptides, in addition to losing the interactions from position 7 due to crystal packing.”

Reviewer #2 (Public review):

Summary:

The authors took a well-characterised (partly by them), important E3 ligase, in the anaphase-promoting complex, and decided to design peptide inhibitors for it based on one of the known interacting motifs (called D-box) from its substrates. They incorporate unnatural amino acids to better occupy the interaction site, improve the binding affinity, and lay foundations for future therapeutics - maybe combining their findings with additional target sites.

Strengths:

The paper is mostly strengths - a logical progression of experiments, very well explained and carried out to a high standard. The authors use a carefully chosen variety of techniques (including X-ray crystallography, multiple binding analyses, and ubiquitination assays) to verify their findings - and they impressively achieve their goals by honing in on tight-binders.

Weaknesses:

Some things are not explained fully and it would be useful to have some clarification. Why did the authors decide to model their inhibitors on the D-box motif and not the other two SLiMs that they describe?

For completeness, in addition to the D-box we did originally construct peptides based on the ABBA and KEN-box motifs, but they did not show any shift in melting temperature of cdc20 in the thermal shift assay whereas the D-box peptides did; consequently, we focused our efforts on the D-box peptides. Moreover, there is much evidence from the literature that points to the unique importance of the D-box motif in mediating productive interactions of substrates with the APC/C (i.e. those leading to polyubiquitination & degradation). One of the clearest examples is a study by Mark Hall’s lab (described in Qin et al. 2016), which tested the degradation of 15 substrates of yeast APC/C in strains carrying alleles of Cdh1 in which the docking sites for D-box, KEN or ABBA were mutated. They observed that whereas degradation of all 15 substrates depended on D-box binding, only a subset required the KEN binding site on Cdh1 and only one required the ABBA binding site. A more recent study from David Morgan’s lab (Hartooni et al. 2022) looking at binding affinities of different degron peptides concluded that KEN motif has very low affinity for Cdc20 and is unlikely to mediate degradation of APC/C-Cdc20 substrates. Engagement of substrate with the D-box receptor is therefore the most critical event mediating APC/C activity and the interaction that needs to be blocked for most effective inhibition of substrate degradation.

We have added the following text to the Results section “Design of D-box peptides” (page 10):

“We focused on D-box peptides, as there is much evidence from the literature that points to the unique importance of the D-box motif in mediating productive interactions of substrates with the APC/C (i.e. those leading to polyubiquitination & degradation). One of the clearest examples is a study that tested the degradation of 15 substrates of yeast APC/C in strains carrying alleles of Cdh1 in which the docking sites for D-box, KEN or ABBA were mutated ((Qin et al. 2017)). They observed that, whereas degradation of all 15 substrates depended on D-box binding, only a subset required the KEN binding site on Cdh1 and only one required the ABBA binding site. A more recent study (Hartooni et al. 2022) of binding affinities of different degron peptides concluded that KEN motif has very low affinity for Cdc20 and is unlikely to mediate degradation of APC/C-Cdc20 substrates. Engagement of substrate with the D-box receptor is therefore the most critical event mediating APC/C activity and the interaction that needs to be blocked for most effective inhibition of substrate degradation.”

What exactly do they mean when they say their 'observation is consistent with the idea that high-affinity binding at degron binding sites on APC/C, such as in the case of the yeast 'pseudo-substrate' inhibitor Acm1, acts to impede polyubiquitination of the bound protein'? It's an interesting thing to think about, and probably the paper they cite explains it more but I would like to know without having to find that other paper.

Interesting results from a number of labs (Choi et al. 2008, Enquist-Newman et al. 2008, Burton et al. 2011, Qin et al. 2019) have shown that mutation of degron SLiMs in Acm1 that weaken interaction with the APC/C have the unexpected consequence of converting Acm1 from APC/C inhibitor to APC/C substrate. A necessary conclusion of these studies is that the outcome of degron binding (i.e. whether the binder functions as substrate or inhibitor) depends on factors other than D-box affinity and that D-box affinity can counteract them. One idea is that if a binder interacts too tightly, this removes some flexibility required for the polyubiquitination process. The most recent study on this question (Qin et al.2019) specifically pins the explanation for the inhibitory function of the high affinity D-box in Acm1 on its ‘D-box Extension’ (i.e. residues 8-12) preventing interaction with APC10. In our current study, the binding affinity of peptides is measured against Cdc20. In cellular assays however, the D-box must also engage APC10 for degradation to occur. It may be that the peptide binding most strongly to the D-box pocket on Cdc20 is less able to bind to APC10 and therefore less effective in triggering APC10-dependent steps in the polyubiquitination pathway. The important Hartooni et al. paper from David Morgan’s lab confirms that even though the binding of D-box residues to APC10 is very weak on its own, it can contribute 100X increase in affinity of a peptide by adding cooperativity to the interaction of D-box with co-activator. Re Figure 6 and the fact that we did look at peptide binding in cells, these experiments were done in unsynchronised cells, so most Cdc20 would not be bound to APC/C.

We have modified the text (page 18) from:

“However, we found the opposite effect: D2 and D3 showed increased rates of mNeon degradation compared to D1 and D19 (Fig. 8C,D). This observation is consistent with the idea that high-affinity binding at degron binding sites on APC/C, such as in the case of the yeast ‘pseudo-substrate’ inhibitor Acm1, acts to impede polyubiquitination of the bound protein (Qin et al. 2019). Indeed, there is no evidence that Hsl1, which is the highest affinity natural D-box (D1) used in our study, is degraded any more rapidly than other substrates of APC/C in yeast mitosis. As shown in Qin et al., mutation of the high affinity D-box in Acm1 converts it from inhibitor to substrate (Qin et al. 2019). Overall, our results support the conclusions that all the D-box peptides engage productively with the APC/C and that the highest affinity interactors act as inhibitors rather than functional degrons of APC/C.”

to:

“However, we found the opposite effect: D2 and D3 showed increased rates of mNeon degradation compared to D1 and D19 (Fig. 8C,D). This observation is consistent with conclusions from other studies that affinity of degron binding does not necessarily correlate with efficiency of degradation. Indeed, there is no evidence that Hsl1, which is the highest affinity natural D-box (D1) used in our study, is degraded any more rapidly than other substrates of APC/C in yeast mitosis. A number of studies of a yeast ‘pseudo-substrate’ inhibitor Acm1, have shown that mutation of the high affinity D-box in Acm1 converts it from inhibitor to substrate (Choi et al. 2008, Enquist-Newman et al. 2008, Burton et al. 2011) through a mechanism that governs recruitment of APC10 (Qin et al. 2019). Our study does not consider the contribution of APC10 to binding of our peptides to APC/CCdc20 complex, but since there is strong cooperativity provided by this additional interaction (Hartooni et al. 2022) we propose this as the critical factor in determining the ability of the different peptides to mediate degradation of associated mNeon.”

Reviewer #3 (Public review):

Summary:

Eapen and coworkers use a rational design approach to generate new peptide-inspired ligands at the D-box interface of cdc20. These new peptides serve as new starting points for blocking APC/C in the context of cancer, as well as manipulating APC/C for targeted protein degradation therapeutic approaches.

Strengths:

The characterization of new peptide-like ligands is generally solid and multifaceted, including binding assays, thermal stability enhancement in vitro and in cells, X-ray crystallography, and degradation assays.

Weaknesses:

One important finding of the study is that the strongest binders did not correlate with the fastest degradation in a cellular assay, but explanations for this behavior were not supported experimentally. Some minor issues regarding experimental replicates and details were also noted.

Interesting results from a number of labs (Choi et al. 2008, Enquist-Newman et al. 2008, Burton et al. 2011, Qin et al. 2019) have shown that mutation of degron SLiMs in Acm1 that weaken interaction with the APC/C have the unexpected consequence of converting Acm1 from APC/C inhibitor to APC/C substrate. A necessary conclusion of these studies is that the outcome of degron binding (i.e. whether the binder functions as substrate or inhibitor) depends on factors other than D-box affinity and that D-box affinity can counteract them. One idea is that if a binder interacts too tightly, this removes some flexibility required for the polyubiquitination process. The most recent study on this question (Qin et al.2019) specifically pins the explanation for the inhibitory function of the high affinity D-box in Acm1 on its ‘D-box Extension’ (i.e. residues 8-12) preventing interaction with APC10. In our current study, the binding affinity of peptides is measured against Cdc20. In cellular assays however, the D-box must also engage APC10 for degradation to occur. It may be that the peptide binding most strongly to the D-box pocket on Cdc20 is less able to bind to APC10 and therefore less effective in triggering APC10-dependent steps in the polyubiquitination pathway. The important Hartooni et al. paper from David Morgan’s lab confirms that even though the binding of D-box residues to APC10 is very weak on its own, it can contribute 100X increase in affinity of a peptide by adding cooperativity to the interaction of D-box with co-activator. Re Figure 6 and the fact that we did look at peptide binding in cells, these experiments were done in unsynchronised cells, so most Cdc20 would not be bound to APC/C.

We have modified the text (page 18) from:

“However, we found the opposite effect: D2 and D3 showed increased rates of mNeon degradation compared to D1 and D19 (Fig. 8C,D). This observation is consistent with the idea that high-affinity binding at degron binding sites on APC/C, such as in the case of the yeast ‘pseudo-substrate’ inhibitor Acm1, acts to impede polyubiquitination of the bound protein (Qin et al. 2019). Indeed, there is no evidence that Hsl1, which is the highest affinity natural D-box (D1) used in our study, is degraded any more rapidly than other substrates of APC/C in yeast mitosis. As shown in Qin et al., mutation of the high affinity D-box in Acm1 converts it from inhibitor to substrate (Qin et al. 2019). Overall, our results support the conclusions that all the D-box peptides engage productively with the APC/C and that the highest affinity interactors act as inhibitors rather than functional degrons of APC/C.”

to:

“However, we found the opposite effect: D2 and D3 showed increased rates of mNeon degradation compared to D1 and D19 (Fig. 8C,D). This observation is consistent with conclusions from other studies that affinity of degron binding does not necessarily correlate with efficiency of degradation. Indeed, there is no evidence that Hsl1, which is the highest affinity natural D-box (D1) used in our study, is degraded any more rapidly than other substrates of APC/C in yeast mitosis. A number of studies of a yeast ‘pseudo-substrate’ inhibitor Acm1, have shown that mutation of the high affinity D-box in Acm1 converts it from inhibitor to substrate (Choi et al. 2008, Enquist-Newman et al. 2008, Burton et al. 2011) through a mechanism that governs recruitment of APC10 (Qin et al. 2019). Our study does not consider the contribution of APC10 to binding of our peptides to APC/CCdc20 complex, but since there is strong cooperativity provided by this additional interaction (Hartooni et al. 2022) we propose this as the critical factor in determining the ability of the different peptides to mediate degradation of associated mNeon.”

Recommendations for the authors:

Reviewer #1 (Recommendations for the authors):

(1) On page 12 (towards the end), the author stated D10 contained an A3P mutation, they meant P3A right? 'To test this hypothesis, we proceeded to synthesise D10, a derivative of D4 containing an A3P single point mutation.'

We thank the reviewer for spotting this typo, which we have corrected.

(2) Have the authors considered other orthogonal approaches to cross-examine/validate binding affinities? That said, I do not think extra experiments are necessary.

We did not explore further orthogonal approaches due to the challenges of producing sufficient amounts of the Cdc20 protein. Due to the low affinities of many peptides for Cdc20, many techniques would have required more protein than we were able to produce. We believe that the qualitative TSA combined with the SPR is sufficient to convince the readers; indeed there is a correlation between SPR-determined binding affinities and the thermal shifts: For the natural amino acid-containing peptides (Table 1) D19 has the highest affinity and causes the largest thermal shift in the Cdc20 melting temperature, D10 has the lowest affinity and causes the smallest thermal shift, and D1, D3, D4, and D5 and all rank in the middle by both techniques. For those peptides containing unnatural amino acids (Table 2), again higher affinities are reflected in larger thermal shifts.

Reviewer #2 (Recommendations for the authors):

The data seem fine to me. I would appreciate a little more detail on the points mentioned in the public review. Also a thorough reread, maybe by a disinterested party as there are various typos that could be corrected - all in all an excellent clear paper that encompasses a lot of work.

A colleague has carefully checked the manuscript, and typos have been corrected.

Version 2

Abstract

E3 ubiquitin ligases engage their substrates via ‘degrons’ - short linear motifs typically located within intrinsically disordered regions of substrates. As these enzymes are large, multi-subunit complexes that generally lack natural small-molecule ligands and are hard to drug via conventional means, alternative strategies are needed to target them in diseases, and peptide-based inhibitors derived from degrons represent a promising approach. Here we explore peptide inhibitors of Cdc20, a substrate-recognition subunit and activator of the E3 ubiquitin ligase the anaphase promoting complex/cyclosome (APC/C) that is essential in mitosis and consequently of interest as an anti-cancer target. APC/C engages substrates via degrons that include the ‘Destruction box’ (D-box) motif. We used a rational design approach to construct binders containing unnatural amino acids aimed at better filling a hydrophobic pocket on the surface of Cdc20. We confirmed binding by thermal-shift assays and surface plasmon resonance and determined the structures of a number of the Cdc20- peptide complexes. Using a cellular thermal shift assay, we confirmed that the D-box peptides also bind to and stabilise Cdc20 in the cell. We found that the D-box peptides inhibit ubiquitination activity of APC/CCdc20 and are more potent than the small molecule inhibitor Apcin. Lastly, these peptides function as portable degrons capable of driving the degradation of a fused fluorescent protein. Interestingly, we find that although inhibitory activity of the peptides correlates with Cdc20-binding affinity, degradation efficacy does not, which may be due to the complex nature of APC/C regulation and effects of degron binding of subunit recruitment and conformational changes. Our study lays the groundwork for the further development of these peptides as molecular therapeutics for blocking APC/C as well as potentially also for harnessing APC/C for targeted protein degradation.

Introduction

E3 ubiquitin ligases recognise their substrates via short linear motifs (SLiMs) known as degrons on the substrates, which are often located in intrinsically disordered regions and bind with rather weak affinities (in the micromolar range) Tompa et al. 2008; Min et al. 2013; Guharoy et al. 2016). Recognition of the substrate by the proteasome recognition depends on the number, type, and length of polyubiquitin chains and requires poly-ubiquitination through multiple rounds of recruitment of a ubiquitin-loaded conjugating E2 enzyme to the substrate-bound E3. There are numerous examples in which ubiquitination and degradation requires the E3 to bind to multiple degrons on the substrate, a particularly striking case being the anaphase promoting complex/cyclosome APC/C, a giant multi-protein complex that controls progression of cells out of mitosis via the coordinated ubiquitination and degradation of over 100 different substrates (Karamysheva et al. 2009; Fiore et al. 2016; Tian et al. 2012; Okoye et al. 2022). APC/C function requires one of two coactivators, Cdc20 and Cdh1/FZR1 (Min et al. 2013; Davey and Morgan 2016; Bakos et al. 2018), which comprise a substrate-binding WD40 domain and N- and C- terminal disordered regions that bind to other APC/C subunits and lead to conformational change and enhanced E2 binding. APC/CCdc20 activity is further controlled by the mitotic checkpoint complex (MCC) that inhibits APC/C in the presence of faulty chromosome attachments to the mitotic spindle (Izawa and Pines 2011; Fiore et al. 2016; Qiao et al. 2016; Alfieri, Zhang, and Barford 2017; Watson et al. 2019). APC/CCdc20 activity reduces as cells exit mitosis, the coactivator is switched to Cdh1/FZR1, and APC/CFZR1 is active until the end of G1 phase. In addition to coactivator switching and post-translational modifications, the APC/C further controls the order of substrate degradation via differential degron-binding affinities(Davey and Morgan 2016; Alfieri, Zhang, and Barford 2017; Bodrug et al. 2021; Okoye et al. 2022) leading to different processivity of ubiquitination and chain linkages and thereby effective engagement of the proteasome, although the precise details of the relationship between APC/C-substrate interactions and the number, type and length of polyubiquitin chains formed have not been fully resolved.

High-resolution structures of the APC/C and its interactions with substrates and E2s have provided tremendous insights into the mechanism of ubiquitination and degradation (Barford 2020). Coactivators Cdc20 and FZR1 have at least three binding sites for substrate degrons: the “Destruction-box” (D-box), the KEN motif, and the ABBA motif (thought to be required for Cyclin A degradation only (Qin et al. 2017). A cryo-EM study of the structure of APC/C-FZR1 in complex with its pseudo-substrate inhibitor Acm1 showed that all three of the Acm1 degrons can bind simultaneously (He et al. 2013). The KEN motif is recognised by the ‘top’ surface of the WD40 beta-propeller domain of the co-activator, and the ABBA motif binds to the ‘side’ of the beta-propeller at one of its blades. The D-box binding site comprises a cleft between two of the beta-propeller blades and the neighbouring APC/C subunit APC10, which is thought to result in stabilization of the active complex (Burton and Tsakraklides 2005; Buschhorn et al. 2010; Da Fonseca et al. 2010; L. Chang et al. 2014; Matyskiela and Morgan 2009; Qin et al. 2019). Most recently, single-molecule studies have shed new insights into the key role of degron multivalency in enabling efficient substrate ubiquitination and degradation (Hartooni et al. 2022).

Structure and Function of APC/C.

(A) Schematic of APC/C activity during mitotic exit, indicating the switch in co-activator from Cdc20 to FZR1. Most substrates contain variable degrons (D-box in green, KEN in yellow) present in IDRs (B) Domain structuring of Cdc20 comprising an N-terminal IDR with the C-box, KEN-box, and CRY-box motifs, the central WD40 domain responsible for substrate recruitment via the degron binding sites and the C-terminal IDR containing the IR-tail. (C) Schematic of the structure of the Cdc20 WD40 domain (PDB: 4GGC) overlaid with those of the WD40 domain in complex with Acm1 D-box and ABBA motif peptides (PDB: 4BH6) (He et al. 2013) and the KEN-box peptide 4GGD) (Tian et al. 2012).

Inhibitors of APC/CCdc20 activity represent an interesting therapeutic approach to target dividing cells in cancer. Given the large size of the APC/C machine (11 subunits) and the complex mechanisms described above that regulate its function, it is not surprising that it is challenging to target. Apcin and TAME are recently identified small-molecule inhibitors, but they have limited activity and complicated output (Richeson et al. 2020; Sackton et al. 2014a). In this paper we use a rational design approach, based on D-box consensus sequences and a ‘Super D-box’ peptide derived from Hsl1, and examination of the Cdc20-degron interface, to design a series of more potent binders containing unnatural amino acids aimed at better filling the hydrophobic pocket on the interaction interface. We quantified binding by thermal shift assays (TSA) and surface plasmon resonance (SPR) and used a cellular thermal shift assay (CETSA) to demonstrate target engagement within the cellular context. The peptides also show functional engagement with APC/C in the cell as evidenced by their ability to drive the degradation of a fluorescent protein. Most strikingly, in vitro ubiquitination assays with recombinant APC/CCdc20 shows that these peptides are more potent inhibitors of Cyclin B1 ubiquitination than Apcin. Interestingly, we find that although inhibitory activity of the peptides correlates with Cdc20-binding affinity, their degradation efficacy does not. This may be due to the complex nature of APC/C degrons and their bipartite interaction with different subunit, role in E2 recruitment, and consequent impact of positioning for effective ubiquitination. The results provide a useful starting point for the further development of these peptides as molecular therapeutics for blocking APC/C as well as potentially also for harnessing APC/C for targeted protein degradation.

Materials and methods

Cloning, expression, and purification of Cdc20

DNA encoding residues 161-477 of human Cdc20 (Cdc20WD40) was cloned into a pU1 vector with an N-terminal His6-tag followed by a TEV protease cleavage site. Plasmid was then transformed into DH10 MultiBac cells expressing a Cre-recombinase. Positive clones were grown up and bacmid DNA prepared by standard protocols. Sf21 or Sf9 cells were grown at 27°C in Erlenmeyer flasks (Corning) and maintained in mid-log phase of growth prior to all experiments. High titre baculovirus was produced by transfecting bacmid DNA into Sf21 cells at 0.5 × 106 cells/ ml cells using Superfect (Qiagen) in 24 deep-well blocks. Virus was harvested 1-week post transfection. For protein over-expression, Sf21 or Sf9 cells were infected with the virus stock and harvested about 60 hours post infection. Cell pellets were resuspended in 50 mM Tris-HCl, 300 mM NaCl, 1 mM MgCl2, 1 mM TCEP, 5% (v/v) glycerol, SigmaFAST EDTA-free protease inhibitor cocktail (1 tablet/100 ml), pH 8.5. Resuspended pellets were lysed by one freeze-thaw cycle at-80°C. Lysates were then clarified by centrifugation at 45,000 × g for 45 minutes at 4°C. Supernatants were flowed over a 5 ml HisTrap Excel column and washed with 20 column volumes (CV) of 50 mM Tris-HCl, 300 mM NaCl, 10 mM imidazole, 1 mM MgCl2, 1 mM TCEP, 5% (v/v) glycerol, pH 8.5. Proteins were then eluted with the above buffer including 300 mM imidazole directly into a 26/10 desalting column pre-equilibrated in the above buffer without imidazole. Eluted protein fractions were then pooled, and the His6-tag was removed using His6- TEV protease (S219V) overnight at 4°C. Proteins were then flowed back over a 5 ml HisTrap Excel column, collecting the flow-through containing the non-tagged Cdc20 protein. Protein eluent was then diluted in 25 mM Tris-HCl buffer, 1 mM MgCl2, 1 mM TCEP, 5% (v/v) glycerol, pH 8.5 to a final concentration of 30 mM NaCl. Proteins were then loaded onto a MonoQ 10/100 GL column and eluted over 20 CV with 1 M NaCl. Protein fractions containing the Cdc20 protein were then pooled and concentrated before separating on a Superdex 75 increase 10/300 GL column in the final buffer containing 25 mM Tris-HCl, 150 mM NaCl, 1 mM MgCl2, 5 mM TCEP, 5% (v/v) glycerol, pH 8.5.

Minimal biotinylation of Cdc20 for SPR

The protocol for minimal biotinylation was adapted from Papalia and Myszka (Papalia and Myszka 2010). Purified Cdc20WD40 was diluted into the reaction buffer (10 mM Tris-HCl, 150 mM NaCl, 1 mM MgCl2, 5% (v/v) glycerol, 1 mM TCEP, pH 7.3). A 0.9:1 molar ratio of Sulfo-NHS-LC- LC-Biotin (Thermo Fisher Scientific, A35358) was added to the diluted Cdc20 protein. The contents were briefly mixed by vortex and incubated on ice for 3 hours. The sample was then separated on a Superdex 75 Increase 10/300 GL column to remove free biotin.

Peptide synthesis and purification

Peptides synthesis was performed on a 0.1 mmol scale using Ramage-ChemMatrix® resin (Sigma Aldrich). Fmoc-L-amino acids (2 eq.), HATU (2 eq.) and HOAt (2 eq.) were dissolved in 2 mL of NMP. DIEA (3.4 eq.) were used to activate the coupling mixture. Activated Fmoc-L-amino acids were coupled for 10 minutes (Fmoc-L-Arginine, 2 × 5 eq., 30 minutes). Resins were washed in DMF and deprotected in 20% piperidine in DMF for 15 minutes. All peptides were N- terminally acetylated in 4 ml DMF, 4 mL acetic anhydride, 2 mL DIEA for 10 minutes. A peptide cleavage cocktail consisting of 93% TFA, 3.5% TIPS and 3.5% ddH2O was used to deprotect and cleave the peptide from the resin for 1 hour. The eluate was triturated by the addition of diethyl ether and the resulting precipitate was isolated by brief centrifugation. All peptides were characterised by LCMS using a Waters ACQUITY H-Class UPLC with an ESCi Multi-Mode Ionisation Waters SQ Detector 2 spectrometer. LC was performed on a ACQUITY UPLC® CSH C18 (2.1 mm × 50 mm, 1.7 µm, 130 Å) at 40°C, with a PDA eλ detector 220 – 800 nm, interval 1.2 nm. The following solvents and gradients were used for LC runs. Solvent A: 2 mM NH4OAc in 95% H2O, 5% MeCN, solvent B: 100% MeCN, solvent C: 2% Formic acid from 5-95% B with a constant of 5% C over 1 minute at 0.6 ml/min. Analytical and semi preparative HPLC runs were performed on an Agilent 1260 Infinity system using a Supelcosil ABZ+PLUS (150 mm × 4.6 mm, 3 µm) and Supelcosil ABZ+PLUS (250 mm × 21.2 mm, 5 µm), respectively. Peptides were eluted with a linear gradient system (solvent A: 0.1% TFA in H2O, solvent B: 0.05% TFA in MeCN) over 15 minutes at 1ml/min and 20 minutes at 20ml/min, respectively. Eluents were monitored by UV absorbance at 220 nm and 254 nm. Analytical data for all peptides are shown in Figure S5.

Thermal-shift assays (TSA)

Assays were performed using a Roche LightCycler 480 I in 96-well plate format. Each well (20 µl) was prepared with 750 nM of purified Cdc20WD40 and varying concentrations of D-box peptides, Apcin or DMSO (vehicle control) in assay buffer; 25 mM Tris-HCl, 150 mM NaCl, 1 mM MgCl2, 5% (v/v) glycerol, 1 mM TCEP, 1% (v/v) DMSO, 5x SYPRO Orange (Thermo Fisher), pH 8.5. Thermal ramps were conducted from 25°C to 95°C at a rate of 0.03°C/sec and data were collected at a frequency of 20 points/°C. An excitation wavelength of 483 ± 35 nm was used to excite SYPRO Orange, and the fluorescence emission was detected at a 568 ± 20 nm. Measurements were performed in triplicate and errors listed are the standard deviation. Melting temperatures were determined by the minima peak of the negative differential in the ‘Tm calling’ analysis within the in-built analysis software.

Surface Plasmon Resonance (SPR) assays

Experiments were performed using a Biacore T200 instrument (GE healthcare) at 15°C. Biotinylated-Cdc20WD40 was immobilised onto a SA biosensor chip (GE healthcare) in running buffer (10 mM HEPES, 150 mM NaCl, 0.1 mM TCEP, 0.05% (v/v) Tween 20 and 1% (v/v) DMSO, pH 7.4) over flow cells 2, 3 and 4 at varying ligand densities. Flow cell 1 was used as a reference cell. Free biotin binding sites were blocked using amine-PEG4-Biotin. Peptides and Apcin analytes were diluted from DMSO stock solutions in running buffer without DMSO and were buffer matched to 1% DMSO. Titrations of each analyte were run over the sensor chip at a flow rate of 30 µl/min. Binding interactions were detected as a change in response units over the reference flow cell and subtracted from a blank buffer injection. Dissociation constants (KD) were calculated by fitting the response units (RU) at steady-state equilibrium generated by the binding of an analyte to Cdc20WD40 against the concentration of analyte using the following equation: where RUanalyte is the response units at equilibrium during a given injection of a concentration of analyte, [analyte]. RUmax is the maximum response produced by the a given analyte, dependant on the RU of immobilised ligand on a given flow cell. KD is the dissociation constant of a given analyte to the ligand. KD values are shown as the average of measurements from the three reference-subtracted flow cells.

Cellular thermal shift assays (CETSA)

Full-length Cdc20 (residues 1-499) with a C-terminal HiBiT tag (GSVSGWRLFKKISGS, Promega) was cloned into a pcDNA3.1(-) vector. HEK 293T cells were cultured in DMEM + 10% FBS (Sigma Aldrich, F7524) at 5% CO2 in a humidified environment. Cells were grown to 70% confluency in T75 flasks prior to transient transfection with 10 µg of Cdc20_HiBiT_pcDNA3.1(-) plasmid with Lipofectamine 2000 (Invitrogen, ThemoFisher Scientific) according to the manufacturers’ protocol. Cells were harvested after 48 hours by trypsinisation and were subsequently washed twice in PBS with repeated centrifugation at 1000 × g for 2 minutes. The pellet was then resuspended in lysis buffer (PBS, 1 × SigmaFAST EDTA-free protease inhibitor tablet (Sigma Aldrich), 2 mM NaVO3, 5 mM NaF, pH 7.4) and freeze-thaw lysed in liquid nitrogen. The lysate was clarified by centrifugation at 20,000 × g, 4°C for 20 minutes and the protein concentration of the supernatant was quantified by BCA (Pierce). Lysates were used at a final concentration of 0.2 mg/ml in lysis buffer. Lysates were aliquoted in 300 µL and were spiked with D-box binding site ligands to a concentration of 100 µM maintaining 1% DMSO. Compounds were incubated on ice for 30 minutes prior to aliquoting further into PCR strip tubes on a PCR block at 4°C. Lysate aliquots were then heated on a second PCR block at the indicated temperatures for 3 minutes prior to returning to 4°C. Heated lysates (5 µl) were then transferred into an AlphaPlate light-grey 384-well plate in quadruplicate by multichannel pipette. Nano-Glo HiBiT lytic detection system (Promega) was diluted as per the manufacturers’ instructions and 5 µl were added to each well by multichannel pipette. Lysis buffer and a non-transfected HEK 293T cell lysate were used as negative controls. Following five minutes of incubation on a plate shaker, the plate was measured using a CLARIOStar microplate reader (BMG Labtech), with the detector set to read at 460 ± 80 nm, the focal height at 10.5 cm and the gain adjusted to 2000. Data were normalised to the unheated sample (4°C) and were fitted using a Boltzmann equation to extract the melting temperature (Tm) (Niesen, Berglund, and Vedadi 2007).

Protein crystallisation

Peptide D21 was added to Cdc20WD40 in a stoichiometric manner and was co-concentrated to 1.9 mg/ml. The resulting complex was crystallised in a 2:1 protein to well solution ratio at 20°C using the sitting-drop vapour-diffusion method with a well solution containing 0.1 M MES pH 6.5, 12% (w/v) PEG 6000, 10% (v/v) MPD for Cdc20WD40-D21 and 0.1 M MES pH 6.5, 14% (w/v) PEG 6000, 10% (v/v) MPD for Cdc20WD40-D20 and Cdc20WD40-D7. Crystals grew to a maximum size after 3 days of incubation. For soaking experiments, crystals were first looped and washed through three drops containing 0.1 M MES pH 6.5, 20% (w/v) PEG 6000 to wash out MPD from the crystal. Crystals were then looped and incubated in a solution containing 0.1 M MES pH 6.5, 20% (w/v) PEG 6000 and 2.5 mM D21 or D20 (5% (v/v) DMSO) or D7 (10% (v/v) DMSO) for four hours. Soaked crystals were cryo-protected in the soak solution supplemented with 10% (v/v) glycerol and were flash frozen in liquid nitrogen.

Data collection and structure determination

Diffraction data were collected on beamline I04 at the Diamond Light Source (Oxford, UK) and processed using autoPROC-STARANISO STARANISO (Vonrhein et al. 2018). Phases were obtained by molecular replacement using the crystal structure of human Cdc20 (PDB ID code 4GGC) as the search model (Tian et al. 2012). Iterative model building and refinements were performed with COOT and BUSTER, respectively (Emsley 2010, Bricogne G. et al) Cdc20-D20 and Cdc20-D21 datasets were first refined using Refmac5 within the CCP4i suite ((Winn et al. 2011; Kovalevskiy et al. 2018) before final refinements using BUSTER. Data collection and structure refinement statistics are summarised in Table S1.

Ubiquitination assays

In vitro ubiquitination experiments were performed using APC/C and Cdc20 purified from insect cells (Zhang et al. 2016). 60 nM APC/C, 30 nM Cdc20, 90 nM UBA1, 300 nM UbcH10, 300 nM Ube2S, 35 mM ubiquitin, 1 mM cyclin B1, 5 mM ATP, 10 mM MgCl2, were mixed in a buffer containing 40 mM HEPES (pH 8.0), 80 mM NaCl, 0.6 mM DTT. The reaction was either performed with the indicated concentrations of peptides or DMSO (Sigma Aldrich) as the vehicle control. The reaction was incubated for 30 min at 23 °C and stopped by the addition of one volume of 2x concentrated NuPAGE LDS loading buffer (Invitrogen).

Protein degradation assays

The pEGFP-N1 vector was modified by swapping the EGFP-coding sequence for mNeon-coding sequence using the AgeI/NotI cloning sites. The Aurora kinase A (AURKA) C-terminal fragment (364-403) containing the non-degron R371xxL motif (D0) together with an extended IDR was amplified by PCR and cloned into the modified vector with BamHI/AgeI sites. Round the horn site-directed mutagenesis was used to generate different D-box variants and validated by DNA sequencing. U2OS cells were cultured in DMEM supplemented with 10% FBS, 200 µM Glutamax-1, 100 U/ml penicillin, 100 µg/ml streptomycin, and 250 ng/ml fungizone (all from ThermoFisher Scientific) at 37°C in humidified atmosphere containing 5% CO2. Plasmids were introduced into U2OS cells by electroporation using the Neon™ Transfection System 10 μL Kit (ThermoFisher Scientific) and cells seeded on eight-well microscopy slides (Ibidi) and recovered for 24 hours. DMEM medium was exchanged for phenol red-free Leibovitz’s L15 (ThermoFisher Scientific), supplemented as above. Time-lapse imaging was conducted at 37°C using a widefield imaging platform composed of Olympus IX83 motorized inverted microscope, Spectra-X multi-channel LED widefield illuminator (Lumencor, Beaverton, OR, USA), Optospin filter wheel (Cairn Research, Faversham, UK), CoolSnap MYO CCD camera (Photometrics, Tucson, AZ, USA), automated XY stage (ASI, Eugene, OR, USA) and climate chamber (Digital Pixel, Brighton, UK), all controlled using Micro-Manager software (Edelstein et al. 2014). Fluorescence and phase contrast images of cells in mitosis were acquired with a 40X objective binned at 2×2 at 2-minute intervals. Fluorescence intensity of mNeon in individual mitotic cells was quantified from 16-bit tiff files using ImageJ, by integrating pixel measurements after subtraction of background fluorescence. Degradation curves were synchronized in silico to anaphase onset to generate average curves for multiple cells in each experimental condition.

Results

Quantification of Cdc20-binding activity of the small molecule Apcin

We first produced Cdc20 protein in sufficient quantities for biophysical analysis and then used the known small molecule binder, Apcin, to test that the purified protein was functional and to benchmark our peptide-binding measurements. As Cdc20 comprises a WD40 domain that binds to the different degrons and is flanked on each end by long intrinsically disordered regions, we made a construct comprising the WD40 domain (residues 161 to 477) with an N-terminal His6-tag and expressed this protein in baculovirus as previously described (Sackton et al. 2014b; Tian et al. 2012). We biotinylated Cdc20 at a single-site, as shown by electrospray-ionisation mass spectrometry (Fig. S1). Using TSA and SPR, we confirmed that the purified Cdc20 was capable of binding to Apcin. The KD obtained by SPR was 420 ± 50 nM (Fig. 2).

Biophysical characterisation of Apcin binding to Cdc20WD40 by TSA and SPR.

(A) Representative examples of thermal unfolding traces of Cdc20 WD40 in the presence of 1% DMSO as the vehicle control or Apcin at concentrations of 25, 50 and 100 µM. (B) Corresponding melting temperatures calculated from derivative plots of the thermal unfolding traces. Mean data from triplicate measurements are shown, with error bars representing standard deviations. (C) Reference-subtracted sensorgrams of biotinylated Cdc20WD40 and Apcin. (D) Binding affinity determination of Apcin to Cdc20WD40 domain by steady-state analysis of the sensorgrams.

Design of D-box peptides

We focused on D-box peptides, as there is much evidence from the literature that points to the unique importance of the D-box motif in mediating productive interactions of substrates with the APC/C (i.e. those leading to polyubiquitination & degradation). One of the clearest examples is a study that tested the degradation of 15 substrates of yeast APC/C in strains carrying alleles of Cdh1 in which the docking sites for D-box, KEN or ABBA were mutated ((Qin et al. 2017)). They observed that, whereas degradation of all 15 substrates depended on D-box binding, only a subset required the KEN binding site on Cdh1 and only one required the ABBA binding site. A more recent study (Hartooni et al. 2022) of binding affinities of different degron peptides concluded that KEN motif has very low affinity for Cdc20 and is unlikely to mediate degradation of APC/C-Cdc20 substrates. Engagement of substrate with the D-box receptor is therefore the most critical event mediating APC/C activity and the interaction that needs to be blocked for most effective inhibition of substrate degradation.

Structures of the D-box-APC/C interactions [(Chao et al. 2012; He et al. 2013)] show that there are three key residues, Arginine 1, Leucine 4 and Asparagine 9 in the RxxLxxxxN D-box motif (Fig. 3A). Leucine 4 contacts a hydrophobic pocket in the co-activator subunit, whereas the ‘tail’ of the D-box degron and its flanking sequence (residues 8-12) contact the APC10 subunit. As a starting point we used two peptides, a 10-residue consensus-like sequence derived from Hsl1 (D1: GRAALSDITN) (Burton and Tsakraklides 2005; Frye et al. 2013; Davey and Morgan 2016), and a 9-residue consensus sequence based on known D-box degrons from APC/C substrates (D2: RLPLGDISN) (He et al. 2013). TSA and SPR showed that D1 binds to Cdc20WD40 with a weak affinity (KD = 18.6 ± 0.2 µM) (Table 1). D2 had no detectable affinity by TSA and was consequently not analysed by SPR. We hypothesised that the apparent lack of binding may be due the low solubility in aqueous buffer of D2 rather than an inability to bind. Based on the Cdh1-Acm1 structure, the sidechain of the amino acid at position 2 is likely to be solvent exposed in the context of Cdc20WD40 (Fig. 3A). Given that we observed a measurable affinity with D1, we chose to incorporate Ala at position 2 in all subsequent peptides.

D-box peptide mutations.

(A) Schematic showing the Acm1 D-box peptide bound to yeast FZR1 homologue Cdh1. R119 of the D-box forms H-bond interactions with D256 and E537 of Cdh1. L122 of the D-box buries into the canonical pocket on the surface of Cdh1 (PDB: 4BH6, He et al. 2013). (B) Melting temperature of Cdc20WD40 in the presence of D-box peptides at 25, 50 and 100 µM concentrations, calculated from derivative plots of the thermal unfolding traces. Mean data from triplicate measurements are shown, with error bars representing standard deviations.

Isoleucine at position 7 and Proline at position 3 of the D-box peptide are optimal for binding

From the consensus sequence, we observed that substrate proteins have approximately equal frequency of Val, Leu and Ile at position 7. Based on the yeast Cdh1-Acm1 X-ray crystal structure (Fig. 3A), this interaction appears atypical of hydrophobic interactions given the largely solvent-exposed nature of the amino acid sidechain. Given the similar structural and physical properties of the three aliphatic sidechains, we compared peptides with each of these three amino acids at position 7 and found that D4 with Ile7 had the highest affinity for Cdc20 (1.5-fold higher than D5 with Leu7 (19.6 ± 0.2 µM and 27 ± 1 µM respectively) (Table 1 and Fig. S2C, D). Interestingly, the shorter hydrocarbon chain of Val in D3 gave the weakest affinity, with a KD determined by SPR at 54.4 ± 0.7 µM.

We next investigated the contribution of Proline versus Alanine at position 3 (Table 1). Like the position 7 residues, Pro and Ala appear in approximately equal distribution to each other among known substrate proteins. In the context of D-box degron binding, modelling of our D4 peptide to the S. cerevisiae Cdh1 structure showed that Pro 3 may form a favourable turn in the D-box peptide backbone to allow the side chain of Leu 4 to adopt its canonical pocket (Fig. 3A). To test this hypothesis, we proceeded to synthesise D10, a derivative of D4 containing a P3A single point mutation. As expected, this mutation was significantly detrimental with an affinity of 70 ± 3 µM by SPR and in parallel a loss of thermal stabilisation by TSA (Fig. 3B). Upon confirming our hypothesis, we synthesised a derivative of D1 containing the A3P point mutation, yielding D19 (RAPLSDITN). This substitution resulted in 3-fold increase in affinity (KD = 5.9 ± 0.1 µM) compared with its parental sequence (Fig. S2F, Table 1).

Unnatural amino acids at position 4 of the D-box peptide result in significantly enhanced binding affinity to Cdc20

The surface topology of Cdc20 is largely flat, making it hard to drug. Nevertheless, in Apcin the tri-chlorinated moiety makes particular use of the Leu 4-binding pocket on Cdc20. Taking inspiration from the small molecule, we explored candidate unnatural amino acids to incorporate into the D-box peptides at position 4. Given that the pocket can accommodate a tri-chlorinated carbon moiety within Apcin, we explored similar moieties to append to our D-box peptides. We incorporated (S)-2-amino-4,4-dimethylpentanoic acid (C3) (Fig. 4A) into the backbone sequences of D4, D10 and D19 replacing Leu at position 4, yielding peptides D7, D12, and D20, respectively (Table 2). As expected, the structure-activity relationship (SAR) held true between all peptides, whereby incorporation of the unnatural amino acid increased the binding affinity over 6-fold versus the respective parental peptide (Table 2). Building on this success, we further explored the commercially available halogenated analog, (S)-2-amino-4,4,4-trifluorobutanoic acid (F3) (Fig. 4A), leading to peptide D21 (Table 2). With the tri-fluorinated group, a further increase in binding affinity was achieved (KD = 520 ± 10 nM), which is similar to that of Apcin.

D-box peptides incorporating unnatural amino acids.

(A) Schematics of the two unnatural amino acids used. (B) Thermal stabilisation of the Cdc20WD40 by the two highest affinity peptides D20 and D21 calculated from derivative plots in TSA. SPR reference-subtracted sensorgrams and binding curves for (C) D20, and (D) D21.

Crystal structures of Cdc20-peptide complexes reveal D-box binding mode

Previous attempts to co-crystallise Cdc20 and securin-derived or cyclin B1-derived D-box peptides by Tian and co-workers were unsuccessful (Tian et al. 2012), which may be due to the low affinity of peptides comprising these sequences. Despite the relatively high affinity of D21 and the approximate 1:1.5 ratio of protein to peptide used in co-crystallisation experiments, crystals were absent of peptide ligands and instead contained the 2-methyl-2,4-pentanediol (MPD) molecule in the Leucine-binding cleft (data not shown), originating from the crystallisation well solution. We therefore adopted a similar protocol to that described by Sackton et al., whereby MPD was ‘washed’ out from the crystal prior to performing a soaking experiment with the desired ligand. We attempted these soaking experiments with our four highest affinity peptides, D21, D20, D7 and D19 (in order of highest affinity to lowest) and were able to observe sufficient ligand density for all but D19.

The crystal structures of Cdc20WD40 in complex with each of the other three D-box peptides (Fig. 5A-C) show that they bound to Cdc20WD40 at the canonical D-box degron binding site, with a largely similar topology to the S. cerevisiae Acm1-Cdh1 structure (Fig. 5D (overlay of D21 with Acm1 D-box). The R1 guanidino group of peptides interacts forms hydrogen bonds with the carboxylic acid side chains of D177 and E465 of Cdc20WD40. The nitrogen backbone atom of the (S)-2-amino-4,4,4-trifluorobutanoic acid/ (S)-2-amino-4,4-dimethylpentanoic acid unnatural amino acids also form a hydrogen bond with the carbonyl of D177. Additionally, the carbonyl of S5 belonging to D21/D20 form a H-bond with the nitrogen backbone atom of D177. Lastly, D6 forms inter-molecular H-bonds with R174. We also observed intra-molecular H-bond between the carbonyl of A2 with the amine of G5/S5, in addition the carbonyl of A2 to the hydroxyl of S5 in D21/D20. Crystal packing of an adjacent asymmetric unit of the WD40 domain likely occludes the assumed binding site for the C-terminal three residues (…ITN-NH2). We therefore assume that this is the reason for the lack of observed density in this region of the peptides D20 and D21 (Fig. S3E and S3F, respectively). We believe that it causes a reduction in binding affinities of all peptides in crystallo, given the evidence from SPR highlighting a role of position 7 in the interaction (Table 1). Interestingly, the observed electron density of the peptide correlates with Cdc20 binding affinity: D21 and D20, having the highest affinities, display the clearest electron density allowing six amino acids to be modeled, whereas D7 shows relatively poor density permitting modelling of only four residues. For D19, the lack of density observed likely reflects its intrinsically weaker affinity compared to the other peptides, in addition to losing the interactions from position 7 due to crystal packing. This hypothesis also correlates to the comments made by Tian et al. in their attempt to co-crystallize securin D-box peptides with Cdc20, in the identical space group (Tian et al. 2012).

Crystal structures of Cdc20-D-box complexes.

X-ray crystal structures of peptides (A) D21, (B) D20 & (C) D7 bound to the canonical D-box binding pocket of Cdc20. Intermolecular hydrogen bonds between peptides and Cdc20 are shown by dashed lines. (D) Structural alignment of D21-bound Cdc20 and Acm1 D-box peptide bound to Cdh1 (PDB: 4BH6 (He et al. 2013)). Peptide backbones align to with an RMSD of 1.007Å. Modelled water molecules have been removed from images for clarity.

D-box peptides bind to Cdc20 in the cellular context

We next investigated whether the four highest-affinity peptides D21, D20, D7, and D19 can bind to Cdc20 in the cellular context using a cellular thermal shift assay (CETSA) (Martinez Molina 2013). Sackton et. al previously demonstrated that Apcin can stabilise endogenous Cdc20 by using an isothermal CETSA method (Sackton et al. 2014b). We were able to reproduce this ligand-induced stabilisation of Cdc20 using the more commonly used temperature gradient approach by densitometric analysis of western blots (Fig S3A). However, due to the low-throughput of the assay we also explored a more high-throughput approach by making use of Promega’s split-luciferase HiBiT tag appended to the C-terminus of full-length Cdc20, based on protocols previously described by Martinez and co-workers (Martinez et al. 2018). Notably, the signal is more sensitive and has a larger range of compared to a western blot, and it removes a significant time-consuming centrifugation step from the workflow. We first confirmed that omitting the centrifugation step did not significantly affect the observed Tm of vehicle control samples (Fig. S4B). To further validate that the transfected Cdc20 is functional, we probed binding of 100 µM Apcin, which gave a Tm of 54.4°C ± 0.6 °C (Fig. S4C). We then explored whether the D-box peptides at a fixed concentration stabilise the Cdc20, and for D7, D20 and D21 we observed increases in the thermal stability of Cdc20 that correlated with their binding affinities as previously determined (Fig. 6 and Table 3). The lowest-affinity peptide, D19, did not result in a significant thermal stabilisation of Cdc20.

D-box peptides bind to full-length HiBiT-tagged Cdc20 in the cellular context.

Representative CETSA data are shown for Cdc20-tranfected HEK293T cell lysates incubated with D-box peptides at a concentration of 100 µM.

D-box peptides inhibit APC/CCdc20 ubiquitination activity

We next assessed whether D21 and D20, the two highest affinity peptides, are able to inhibit APC/CCdc20 activity. In the context of Cyclin B1 ubiquitination, we found that both peptides are more potent inhibitors compared with Apcin at the same concentration despite having slightly lower Cdc20-binding affinities than Apcin (Fig. 7).

Inhibition of APC/CCdc20-mediated ubiquitination of Cyclin B1 by D-box peptides and Apcin.

In vitro ubiquitination assays using reconstituted APC/CCdc20 with Cyclin B1 as the substrate for ubiquitination. Lead peptides and Apcin were titrated from 300 µM to 3 µM and showed concentration-dependent inhibition of Cyclin B1 ubiquitination compared to the vehicle control (0.7% DMSO).

D-box peptides are able to target mNeon for degradation

To probe the functionality of the D-box variants at the cellular level, we conducted live cell degradation assays using mNeon fusions containing those peptide sequences that contain only natural amino acids: D1, D2, D3, and D19 (Fig. 8A). The D-box sequences were swapped into an RxxL motif previously shown to have no degron activity (Abdelbaki et al. 2022) which we here refer to as ‘D0’, adjacent to the endogenous C-terminal IDR of AURKA to enable processing of the ubiquitinated fusion proteins at the 26S proteasome. We found that all four new D-box variants tested could target mNeon for degradation, with timing consistent with targeting by APC/CCdc20 (Fig. 8B). We predicted that the higher affinity D-box peptides from the in vitro assays (D1 and D19) would mediate increased rates and extent of degradation compared to the lower affinity peptides (D2 and D3). However, we found the opposite effect: D2 and D3 showed increased rates of mNeon degradation compared to D1 and D19 (Fig. 8C,D). This observation is consistent with conclusions from other studies that affinity of degron binding does not necessarily correlate with efficiency of degradation. Indeed, there is no evidence that Hsl1, which is the highest affinity natural D-box (D1) used in our study, is degraded any more rapidly than other substrates of APC/C in yeast mitosis. A number of studies of a yeast ‘pseudo-substrate’ inhibitor Acm1, have shown that mutation of the high affinity D-box in Acm1 converts it from inhibitor to substrate (Choi et al. 2008; Enquist-Newman, Sullivan, and Morgan 2008; Burton, Xiong, and Solomon 2011) through a mechanism that governs recruitment of APC10 (Qin et al. 2019). Our study does not consider the contribution of APC10 to binding of our peptides to APC/CCdc20 complex, but since there is strong cooperativity provided by this additional interaction (Hartooni et al. 2022) we propose this as the critical factor in determining the ability of the different peptides to mediate degradation of associated mNeon.

D-box variants can drive degradation in mitotic cells.

(A) Schematic of D-box-mNeon constructs used in fluorescence timelapse imaging. (B) mNeon fluorescence levels in individual cells plotted over time to show D-box mediated degradation of mNeon in mitosis. Fluorescence measurements from individual cells are normalized to fluorescence at metaphase then in silico synchronized to anaphase onset. Mean degradation curves are shown, with error bars representing SDs. (C) Degradation rate curves show rate of change in relative fluorescence of the D-box variants and reveal maximum degradation rate for each construct. Error bars are depicted as shaded regions and indicate SDs. (D) Levels of relative fluorescence in each cell at t = 1 hour after anaphase onset. Degradation of each D-box construct was significant relative to D0 control, using Welch’s t-test. ****, p ≤ 0.0001. In (B)-(D), n = D0 (20) D1 (23), D2 (40), D3 (38), D19 (34) with data pooled from two or more independent experiments.

Discussion

Here we quantified D-box peptide binding to Cdc20 and show that binding affinities can be enhanced by incorporating unnatural amino acids to better fill the hydrophobic pockets on the Cdc20 surface. We confirmed the success of this approach by determining X-ray crystal structures of Cdc20-peptide complexes. We showed target engagement by the peptides in the cellular context, and we found that the two highest affinity peptides were more potent inhibitors of APC/CCdc20 activity than the small molecule Apcin. Lastly, we found that the D-box peptide is a portable motif that can drive productive ubiquitination leading to degradation when fused to a fluorescent protein target.

The finding that the peptides were more potent than Apcin as APC/CCdc20 inhibitors was somewhat surprising, since Apcin has a slightly higher Cdc20-binding affinity than the peptides. It suggests that inhibiting APC/CCdc20 ubiquitination activity may require larger molecules to compete with substrates effectively. It may also be that, unlike Apcin, the peptides not only block the interaction of substrates with Cdc20 but additionally the interaction with APC10 and/or prevent the conformational change in APC/C that enables recruitment of the E2. In fact, the mechanism of inhibition by Apcin and D-box peptides could be different – it may be that Apcin-bound Cdc20 can still bind to APC/C, but peptide-bound Cdc20 cannot.

Interestingly, although the inhibitory activity of the D-box peptides roughly correlates with the binding affinity, binding and degradation may be inversely correlated. In addition to the binding of D-box substrates to the co-activators Cdc20 and Cdh1, Qin et al have described how residues C-terminal of the D-box sequence, the ‘D-Box Extension’ DBE motif, influence recruitment of APC10 and potentially APC10 conformational changes enabling the recruitment of the E2 Ube2S (Qin et al. 2019). Ube2S is essential for adding K11 chains, and we showed previously that degradation of all substrates is dramatically slowed down by a lack of Ube2S (Min et al. 2015). The mNeon-D-box constructs used in our current study all contain the same DBE motif, so a potential contribution from this motif will not affect the interpretation of our results, but it could certainly be added as an element in future inhibitor design.

In summary, the finding presented here represent a useful starting point for the further development of APC/C inhibitors as both research tools and also molecular therapeutics. Future directions could involve enhancing potency through avidity by incorporating multiple degrons into our molecules and additions to the D-box core sequence to include motifs that engage other components of the APC/C machinery - namely APC10 and the E2 - thereby not only blocking substrate binding more effectively but also better impeding ubiquitination activity. The results also have implications for the design of small-molecule and peptide-based degraders that harness the APC/C.